Seagrass mapping and monitoring along the coasts of Crete, Greece

### 2011-01-30 MScDiplom_Lemenkova.tex
added 'List of Acronyms', 'Document Outline'

\documentclass[10pt, a4paper]{article}
\usepackage{color}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{calc}
\usepackage{amsmath}
\usepackage{longtable}
\usepackage[svgnames,table]{xcolor}
\usepackage[subfigure]{tocloft}
%\usepackage[table]{xcolor}
\usepackage{paralist}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{fancyheadings}
\topmargin = 1pt
\headheight=10mm
\headsep=7pt
%\ensuremath
\usepackage{lastpage}
\usepackage{float}
\usepackage{subfig}
\usepackage{caption}
\usepackage{chngcntr}
\counterwithin{figure}{section}
\counterwithin{table}{section}
\usepackage{appendix}
\captionsetup{margin=10pt,font=small,labelfont=bf}
\usepackage{wrapfig}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\setcounter{tocdepth}{4}
\parindent 0pt
%\floatstyle{boxed} 
\restylefloat{figure}
\usepackage[a4paper=true, ]{hyperref}

@@

\section*{Abstract}
	\markright{Abstract}
	\addcontentsline{toc}{section}{Abstract}
\pagebreak

\section*{List of Tools}\label{sec:0}
	\markright{List of Tools}
	\addcontentsline{toc}{section}{List of Tools}

@@

\pagenumbering{arabic}
\rfoot{MSc Thesis\\ \emph{"Seagrass mapping and monitoring along the coasts of Crete, Greece"}}
\section{Introduction}
%%%%%%%%%%%%%%%%%%  Section 1   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\counterwithin{figure}{section}
\counterwithin{table}{section}

@@

For our purpose we have chosen the \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{\textit{WASI Water Color Simulator}} RTM software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface enabling us to adjust various environmental parameters.

@@

\subsection[Application of the remote sensing data]{Application of the remote sensing data towards seagrass mapping}
Various methods and approaches have been applied towards mapping of the seagrasses, based on
digitized aerial photographs, GPS data, remote sensing and SCUBA-based fieldwork measurements.
SCUBA-based \textit{in-situ} observations, though providing high resolution and accuracy results in seagrass
mapping, is limited in application, because of their time consumption, weather-dependency and
unsuitability for the case of monitoring large areas of water for small-scale mapping. 
The underwater videography with a GPS is a tool of seagrass monitoring which has certain advantages, i.e.
high spatial and visual resolution, non-destructive sampling, effectiveness at all depths and rapid data
collection in the field \cite{Schultz08}\label{Schultz08}. However, it cannot cover large areas for small-scale mapping.
Remote sensing techniques offer clear advantages over other methods of \textit{in-situ} field measurements
and seagrass observations, mentioned above. Preference of the remote sensing methods consists in
their weather-independence, cost-effectiveness, accuracy and spatial coverage, which enables
periodic monitoring of the seagrass meadows and gives access to the distant and unapproachable
areas. Integrated together with GIS vector layers and maps, remote sensing data enable historical
mapping \cite{Carter08,Ardizzone06}\label{Carter08} \label{Ardizzone06} and assessment of change detection.

@@

%%%%%%%%%%%%%%%%%  Section  3  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and methods}
\counterwithin{figure}{section}
\counterwithin{table}{section}

@@

%%%%%%%%%%%%%%%%%  Section  4  %%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\markright{IV. Results}
\counterwithin{figure}{section}
\counterwithin{table}{section}

@@

\pagebreak
\section{Discussion and conclusions}
The suggested methodology of the analysis of seagrass adaptation towards different environments of
the coastal zones can be applied to other marine coastal areas with growing seagrass.\\
More than 50 \% of the world population lives within one km of the coast which results in continued
anthropogenic pressures in coastal regions. \\Therefore, management of coastal resources and shelf
zone protection in the Mediterranean Sea become increasingly important and require large-scale
monitoring and mapping of the shelf zone as vital instrument for the environmental assessment.
\pagebreak

@@

\begin{thebibliography}{999}
\addcontentsline{toc}{chapter}{\textbf{Bibliography}}
\bibliographystyle{plainnat}
\pagenumbering{roman}
\markright{Bibliography}

>>>>

\documentclass[10pt, a4paper]{article}
\usepackage{color}
\usepackage{setspace}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{calc}
\usepackage{amsmath}
\usepackage{longtable}
\usepackage[svgnames,table]{xcolor}
\usepackage[subfigure]{tocloft}
%\usepackage[table]{xcolor}
\usepackage{paralist}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{fancyheadings}
\topmargin = 1pt
\headheight=10mm
\headsep=7pt
%\ensuremath
\usepackage{lastpage}
\usepackage{float}
\usepackage{subfig}
\usepackage{caption}
\usepackage{chngcntr}
\counterwithin{figure}{section}
\counterwithin{table}{section}
\usepackage{appendix}
\captionsetup{margin=10pt,font=small,labelfont=bf}
\usepackage{wrapfig}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\setcounter{tocdepth}{4}
\parindent 0pt
%\floatstyle{boxed} 
\restylefloat{figure}
\usepackage[a4paper=true, ]{hyperref}

@@

\section*{Abstract}
	\markright{Summary Abstract of the MSc Thesis}
	\addcontentsline{toc}{section}{Abstract}
\pagebreak

\section*{Document Outline}
	\markright{Structure of the MSc Thesis}
	\addcontentsline{toc}{section}{Document Outline}
\textit{Chapter 1}: \nameref{sec:1} is an introductory section. It describes general background of the research problem, outlines the need and actuality of the problem (including the environmental vulnerability of the seagrasses), and highlights limitations and possibilities of the remote sensing application for the current work. In this chapter we also set up research objectives, put research questions for hypothesis testing and sketch research approach for the proposed work.\\
In \textit{Chapter 2}: \nameref{sec:2} we review the existing literature and reported research experience on the similar problem: studies of spectral separability of various seafloor types, limitations and advantages of the remote sensing techniques applied for seagrass mapping seagrass ecology and mapping seagrass environment. We considered not only the Mediterranean environment, but also papers from Australian and Chinese scientists, because seagrass monitoring is most actively developed in the southern regions of our planet. The review of the existing RTM algorithms for the retrieval of the optical parameters, as well as description of various tools for the spectrometric measurements - are given in the same Chapter as well. More close attention has been given to the imaging hyperspectral radiometer Trios RAMSES, used for the data collection.\\
\textit{Chapter 3}: \nameref{sec:3} deals with the methods and materials used in current work. We start our discussion from the fieldwork area location and describe tools and instruments used during the fieldwork, as well as sampling design. Procedures of the data pre-processing and capturing imagery from the Google Earth are also presented in the same chapter. \\
\textit{Chapter 4}: \nameref{sec:4} presents to the reader the main results of the current MSc work, obtained during the processing fieldwork and other collected data. It starts from the review of the collected data from different sources, then describes the modelling component of the research, namely WASI colour simulation and plotted resulting graphs. We also discuss here the particularities of various sensors, suitable enough for deriving radiometric information to discriminate sand from the seagrass. Furthermore, this chapter includes analysis of the spectral signatures of various seafloor cover types in conditions of changing environment, possibility and limitations for \textit{P.oceanica} spectral discrimination, statistical analysis of the data sets and mapping based on images processing.\\
\textit{Chapter 5}: \nameref{sec:5} summarizes and briefly discusses again the main results, reported in detail in a previous chapter. This chapter proposes some final discussions to the reader, and comes to the main research outcomes.\\
Finally, we suggest some recommendations for any further research focusing on the seagrass environment along the coasts of Crete (or other Mediterranean areas) in the final \textit{Chapter 6}: \nameref{sec:6}.\\
A lot of plotted graphs, statistical outcomes, tables, auxiliary (yet relevant to our research) pictures and images, illustrating our work - are collected in the \nameref{app}. Due to the standard editorial limitations of the current MSc work, it was not possible to include all them to the main chapters. However, we tried to make the structure of the Appendices most clear and easy-readable as possible, by dividing the Appendices into 10 various sub-sections and referencing to them from the main text where necessary. \\
The \nameref{bib} section includes literature and internet resources used for the current work. The cross-referencing and web hyper-referencing are used in the whole document to make reading more quick, effective and informative.

\section*{List of Acronyms}
	\markright{Acronyms}
	\addcontentsline{toc}{section}{List of Acronyms}
\onehalfspace

\textbf{4AOP}: Automatized Atmospheric Absorption Atlas Operational Release \\
\textbf{6SV1}:  Second Simulation of a Satellite Signal in the Solar Spectrum, Version1\\
\textbf{ANOVA}: ANalysis of VAriance\\
\textbf{ASTER}: Advanced Spaceborne Thermal Emission and Reflection Radiometer\\
\textbf{BRDF}:  Bidirectional Reflectance Distribution Function\\
\textbf{CASI}:		Compact Airborne Spectrographic Imager\\
\textbf{CNES}: 		Centre National d'Etudes Spatiales (of France)\\
\textbf{CRTM}:		Community Radiative Transfer Model\\
\textbf{CZCS}:		 Coastal Zone Color Scanner\\
\textbf{DISORT}:	 Discrete Ordinates Radiative Transfer Program for a Multi-Layered Plane-Parallel Medium\\
\textbf{ENVISAT}:	ENVIronmental SATellite\\
\textbf{ESA}:		European Space Agency\\
\textbf{ERS}: 		European Remote Sensing satellite\\
\textbf{GIS}:		Geographic Information System\\
\textbf{GNU}:		GNU's Not Unix \\
\textbf{GPS}:		 Global Positioning System\\
\textbf{GRASS}:	 Geographic Resources Analysis Support System\\
\textbf{GRETL}:		GNU Regression, Econometrics and Time-series Library\\
\textbf{HyMap}:	Hyperspectral Mapper\\
\textbf{HyperOCR}:	Hyperspectral Ocean Colour Radiometer\\
\textbf{IKONOS}: 	derived its name from the Greek term \textit{eikon} for image\\
\textbf{ILWIS}:		Integrated Land and Water Information System\\
\textbf{iPAQ}:	internet Pocket PC by CompAQ\\
\textbf{KOPRA}:	 Karlsruhe Optimized and Precise Radiative transfer Algorithm\\
\textbf{LaTeX}:		Lamport Tau Epsilon Chi (Greek for \textit{techne}), modified by L.Lamport\\
\textbf{LAI}:	 Leaf Area Index\\
\textbf{Landsat TM}:	 Landsat Thematic Mapper \\
\textbf{Landsat ETM+}:	Landsat Enhanced Thematic Mapper Plus\\
\textbf{LyX}: 		LyriX-files\\
\textbf{MERIS}:		MEdium Resolution Imaging Spectrometer\\
\textbf{MIPAS}:		Michelson Interferometer for Passive Atmospheric Sounding\\
\textbf{MODIS}:	Moderate Resolution Imaging Spectroradiometer\\
\textbf{NASA}:		National Aeronautics and Space Administration\\
\textbf{NIR}:		 Near-Infra-Red\\
\textbf{NOAA}:		 National Oceanic and Atmospheric Administration\\
\textbf{OBC}:		On Board Calibration\\
\textbf{Olympus ST}: 	 Olympus Stylus Tough\\
\textbf{Open EV}: 	Open Earth View\\
\textbf{P.oceanica}:	Posidonia oceanica\\
\textbf{RAMSES}:	RAdiance Measurement SEnSor\\
\textbf{RRTM}:		Rapid Radiative Transfer Model\\
\textbf{RTTOV}: 	Radiative Transfer for TIROS Operational Vertical Sounder\\
\textbf{RTM}:		Radiative Transfer Model\\
\textbf{SAM}:		Scanning Acoustic Microscope\\
\textbf{SCUBA}:	Self Contained Underwater Breathing Apparatus\\
\textbf{SeaWiFS}: 	Sea-viewing Wide Field-of-view Sensor\\
\textbf{SPOT}:		Satellite Pour l'Observation de la Terre\\
\textbf{SPSS	}:	Statistical Package for the Social Sciences\\
\textbf{STDEV}:		Standard Deviation\\
\textbf{SWIR}: 		Short Wave InfraRed\\
\textbf{TIR}: 		Thermal InfraRed\\
\textbf{TIROS}: 	Television and InfraRed Observation Satellite\\
\textbf{UML}:		Unified Modelling Language\\
\textbf{UV}:		UltraViolet (10 nm to 400 nm)\\
\textbf{UVA}:		UltraViolet A, or long waves (315 nm-400 nm)\\
\textbf{UVB}:		 UltraViolet B, or medium waves (280 nm-315 nm)\\
\textbf{VIS}:		VISible\\
\textbf{VNIR}:		Visible and Near-InfraRed\\
\textbf{WASI}:		WAter colour SImulator\\

\pagebreak
\singlespace
\section*{List of Tools}\label{sec:0}
	\markright{List of Tools}
	\addcontentsline{toc}{section}{List of Tools}

@@

\pagenumbering{arabic}
\rfoot{MSc Thesis\\ \emph{"Seagrass mapping and monitoring along the coasts of Crete, Greece"}}
\section{Introduction}\label{sec:1}
%%%%%%%%%%%%%%%%%%  Section 1   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\counterwithin{figure}{section}
\counterwithin{table}{section}

@@

For our purpose we have chosen the \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{\textit{WASI Water Color Simulator}} RTM software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface, enabling to adjust various environmental parameters.

@@

\subsection[Application of the remote sensing...]{Application of the remote sensing data towards seagrass mapping}
Various methods and approaches have been applied towards mapping of the seagrasses, based on
digitized aerial photographs, GPS data, remote sensing and SCUBA-based fieldwork measurements.
SCUBA-based \textit{in-situ} observations, though providing high resolution and accuracy results in seagrass
mapping, is limited in application, because of their time consumption, weather-dependency and
unsuitability for the case of monitoring large areas of water for small-scale mapping. 
The underwater videography with a GPS is a tool of seagrass monitoring which has certain advantages, i.e.
high spatial and visual resolution, non-destructive sampling, effectiveness at all depths and rapid data
collection in the field \cite{Schultz08}\label{Schultz08}. However, it cannot cover large areas for small-scale mapping.
Remote sensing techniques offer clear advantages over other methods of \textit{in-situ} field measurements
and seagrass observations, mentioned above. Preference of the remote sensing methods consists in
their weather-independence, cost-effectiveness, accuracy and spatial coverage, which enables
periodic monitoring of the seagrass meadows and gives access to the distant and unapproachable
areas. Integrated together with GIS vector layers and maps, remote sensing data enable historical
mapping \cite{Carter08,Ardizzone06}\label{Carter08} \label{Ardizzone06} and assessment of change detection.

@@

%%%%%%%%%%%%%%%%%  Section  3  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and methods}\label{sec:3}
\counterwithin{figure}{section}
\counterwithin{table}{section}

@@

%%%%%%%%%%%%%%%%%  Section  4  %%%%%%%%%%%%%%%%%%%%%%%
\section{Results}\label{sec:4}
\markright{IV. Results}
\counterwithin{figure}{section}
\counterwithin{table}{section}

@@

\pagebreak
\section{Discussion and conclusions}\label{sec:5}
The suggested methodology of the analysis of seagrass adaptation towards different environments of
the coastal zones can be applied to other marine coastal areas with growing seagrass.\\
More than 50 \% of the world population lives within one km of the coast which results in continued
anthropogenic pressures in coastal regions. \\Therefore, management of coastal resources and shelf
zone protection in the Mediterranean Sea become increasingly important and require large-scale
monitoring and mapping of the shelf zone as vital instrument for the environmental assessment.
\pagebreak

\section{Recommendations}\label{sec:6}

@@

\begin{thebibliography}{999}
\addcontentsline{toc}{chapter}{\textbf{Bibliography}}\label{bib}
\bibliographystyle{plainnat}
\pagenumbering{roman}
\markright{Bibliography}
